---
title: "R Notebook"
output: html_notebook
---

# Import libraries

```{r}
library(tidyverse)
library(knitr)
library(scales)
library(xtable)
library(ggrepel)
```

# Import raw data

```{r}
d.raw = read_csv('../data/norming/dataFromMongo.csv') 
```

Check how many people failed catch trials

```{r results='asis'}
d.raw %>%
  group_by(aID) %>%
  unite(trial_chunk, trial_type, condition) %>%
  filter(trial_index == last(trial_index)) %>%
  group_by(trial_chunk) %>%
  tally() %>%
  kable() %>%
  cat(sep = '\n')
```

Only include participants who completed exactly 89 color-picker trials (all 80 words + 9 practice/catch)

```{r}
complete_games = d.raw %>%
  filter(trial_type == 'color-picker') %>%
  group_by(aID) %>%
  tally() %>%
  filter(n == 89) %>% 
  pull(aID)

cat('including', length(complete_games), 'complete games')
```



```{r}
d <- d.raw %>%
  filter(aID %in% complete_games)

d %>%
  group_by(wordSetID, aID) %>%
  tally() %>%
  group_by(wordSetID) %>%
  tally()
```

Save filtered data for Python scripts

```{r}
words <- d %>%
  pull(word) %>%
  unique()

colorAssociations <- d %>%
  filter(trial_type == 'color-picker')

write.csv(colorAssociations, "../data/norming/colorPickerData-0.csv", row.names = TRUE)
  
```

Visualize color grids

```{r}
d %>%
  filter(trial_type == 'color-picker') %>%
  mutate(hex = rgb(response_r/255, response_g/255, response_b/255)) %>%
  group_by(word) %>%
  do(ggsave(.$word, device = 'png', plot = show_col(.$hex, labels = F)))
```

Relationship between variance and entropy

```{r}
read_csv('../data/norming/compiled-variance-entropy-glasgowRatings.csv') %>%
  filter(imageability != 'N/A') %>%
  mutate(imageability = as.numeric(imageability),
         concreteness = as.numeric(concreteness)) %>%
  ggplot(aes(x = variance, y = entropy)) +
    geom_text(aes(label = word)) +
    geom_smooth(method ='lm')
```


```{r}
d = read_csv('../data/norming/compiled-variance-entropy-glasgowRatings.csv') 
```

```{r}
filteredData = d %>%
  filter(imageability != 'N/A') %>%
  mutate(imageability = as.numeric(imageability),
         concreteness = as.numeric(concreteness))
```
# visualize variables

```{r}
filteredData %>%
  mutate(log_variance = log(variance)) %>%
  pivot_longer(-word) %>%
  ggplot(aes(x = value)) +
    geom_histogram() +
    facet_wrap(~ name, scales='free') +
    theme_few()
```

```{r}
scatter.smooth(x=filteredData$variance, y=filteredData$imageability)
scatter.smooth(x=filteredData$variance, y=filteredData$concreteness)
scatter.smooth(x=filteredData$entropy, y=filteredData$imageability)
scatter.smooth(x=filteredData$entropy, y=filteredData$concreteness)
```

# Examine relationship between DVs

```{r}
cor(filteredData$entropy, filteredData$variance, method = 'spearman')
cor(filteredData$entropy, filteredData$concreteness, method = 'spearman')
cor(filteredData$entropy, filteredData$imageability, method = 'spearman')
```

# Build and diagnose linear model for IMAGEABILITY

```{r}
# correlation
cat('variance-imageability correlation = ', cor(filteredData$variance, filteredData$imageability))
cor.test(filteredData$variance, filteredData$imageability)

# get summary stats
linearModIMAG <- lm(log(variance) ~ imageability, data=filteredData) # imageability as a function of variance
imagSummary <- summary(linearModIMAG) # capture model summary as an object
print(imagSummary)

# calculate summary stats for practice
imagCoeffs <- imagSummary$coefficients # model coefficients
beta.estimate <- imagCoeffs["variance", "Estimate"]  # get beta estimate for variance
std.error <- imagCoeffs["variance", "Std. Error"]  # get std.error for variance

t_value_imag <- beta.estimate/std.error  # calc t statistic
# larger t-value indicates that it is less likely that the coefficient is not equal to zero purely by chance

p_value_imag <- 2*pt(-abs(t_value), df=nrow(filteredData)-ncol(filteredData))  # calc p Value
# when p Value is less than significance level (< 0.05), we can safely reject the null hypothesis that the co-efficient β of the predictor is zero

f_statistic <- linearModIMAG$fstatistic[1]  # fstatistic
f <- summary(linearModIMAG)$fstatistic  # parameters for model p-value calc
model_p <- pf(f[1], f[2], f[3], lower=FALSE)
```

# Build and diagnose linear model for CONCRETENESS
```{r}
# correlation
cat('entropy-concreteness correlation = ', cor(filteredData$entropy, filteredData$concreteness))
cat('concreteness-imagability correlation = ', cor(filteredData$imageability, filteredData$concreteness))
cor.test(filteredData$entropy, filteredData$concreteness)

# get summary stats
# concreteness as a function of variance
linearModImg <- filteredData %>%
  lm(log(entropy) ~ imageability,
     data = .)

linearModConcrete <- filteredData %>%
  lm(log(entropy) ~ concreteness, 
     data=.) 

linearModBoth <- filteredData %>%
  lm(log(entropy) ~ imageability + concreteness,
     data=.) 

anova(linearModImg, linearModBoth)
anova(linearModConcrete, linearModBoth)
cncSummary <- summary(linearModBoth)
print(cncSummary)
```

```{r}
linearMod <- lm(variance ~ imageability + concreteness, data=filteredData)
print(summary(linearMod))

linearMod <- lm(entropy ~ imageability + concreteness, data=filteredData)
print(summary(linearMod))
```



# Attempt at mixed effects model
http://www.bodowinter.com/uploads/1/2/9/3/129362560/bw_lme_tutorial2.pdf
```{r}
library(lme4)

mixedModel <- lmer(imageability ~ variance + (1|word), data=filteredData, control=lmerControl(check.nobs.vs.nlev = "ignore", check.nobs.vs.rankZ = "ignore", check.nobs.vs.nRE="ignore"))
summary(mixedModel)
```

```{r}
ordered <- filteredData[order( filteredData[,3] ),]
barplot(ordered$entropy, main="Entropy", xlab="words")
```

# Analysis 3: Compare expectations with analyses

---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(knitr)
library(scales)
library(xtable)
library(broom)
```
# load data
```{r}
dExpectation = read_csv('./expectation/expectations-by-condition.csv') 
dTruth = read_csv('./expectation/ground-truth-by-condition.csv') 
```

# word level response
```{r}
# Average expectation rating slider 
avgExpectation <- aggregate(dExpectation[, 2], list(dExpectation$word), mean)
# Each person and say what percent matched their response -> then average that 
avgTruth <- aggregate(dTruth[, 2], list(dTruth$word), mean)

# Z score averages of expectation rating and actual percent matches (across participants) and compare
expectation <- aggregate(dExpectation[, 2], list(dExpectation$word), mean) %>% 
  mutate(zscore = (sliderResponse - mean(sliderResponse))/sd(sliderResponse))

truth <-  aggregate(dTruth[, 2], list(dTruth$word), mean) %>% 
  mutate(zscore = (groundTruth - mean(groundTruth))/sd(groundTruth))
```
# plot word level responses (averaged over participants + z-scored)
```{r}
scatter.smooth(x=expectation$zscore, y=truth$zscore, xlab="Average expecation rating", ylab="Average percent agreement", main="Expected agreement vs. actual agreement")
```

# Expectation ratings box plot
```{r}
library("ggpubr")
ggboxplot(dExpectation, y = "sliderResponse", main= "Expected Agreement by Condition",
          color = "condition", palette = c("#D8644B", "#4B7CD8"),
          order = c("concrete", "abstract"), ylim = c(0, 100),
          ylab = "% of population expected to share one's color association", xlab="words")
```

# Ground truth boxplot
```{r}


ggboxplot(dTruth, y = "groundTruth", main= "Actual Agreement by Condition",
          color = "condition", palette = c("#D8644B", "#4B7CD8"),
          order = c("concrete", "abstract"), ylim = c(0, 100),
          ylab = "% of population actually sharing color association", xlab="words")
```

# plotting expectation for top and bottom 10 words individually
```{r}
selectedWords <- filter(dExpectation, word %in% c("lemon", "sun", "tomato", "fire", "apple", "spinach", "vegetable", "tree", "daffodil"))

boxplot(sliderResponse~word, data=selectedWords, main="Expected Agreement",
   xlab="Words", ylab="% of population expected to share one's color association", ylim = c(0, 100))
```

# plotting actual agreement for top 10 words individually
```{r}
selectedWords <- filter(dTruth, word %in% c("lemon", "sun", "tomato", "fire", "apple", "spinach", "vegetable", "tree", "daffodil"))

boxplot(groundTruth~word, data=selectedWords, main="Actual Agreement",
   xlab="Words", ylab="% of population actually sharing association", ylim = c(0, 100))
```


<!-- # compare expectation to entropy for each word -->
<!-- ```{r} -->

<!-- filtered <- d %>% -->
<!--   mutate(sliderResponse = as.numeric(sliderResponse)) -->
<!-- # calculate mean per group  -->
<!-- avgExpectation <- aggregate(filtered$sliderResponse, by = list(filtered$word), FUN = mean) -->

<!-- # load and format entropy data -->
<!-- f = read_csv('./entropy/sorted-entropies-all-both.csv')  -->

<!-- f = f %>% -->
<!--   mutate(both = as.numeric(both)) -

<!-- scatter.smooth(x=f$both, y=avgExpectation$x, xlab = "Entropy of participants' color associations for all words", ylab = "Average % of population expected to share one's color association") -->
<!-- ``` -->

<!-- # attempt to combine entropy ratings and slider responses to do a mixed effect model with subject as the random effect -->
<!-- ```{r} -->
<!-- library(lme4) -->

<!-- # NEED TO JOIN DATAFRAMES -->
<!-- mixedModel <- lmer(filtered$entropy ~ d$sliderResponse  + (1|d$aID)) -->
<!-- summary(mixedModel) -->
<!-- ``` -->


